\bf Visual diagnostics for constrained optimisation with application to guided tours

### 2020-10-07 paper/paper.tex
fixing on section 5 examples

\begin{figure}
\includegraphics[width=1\linewidth,height=0.6\textheight]{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figures/tour_path_keynote/tour_path_keynote.001} \caption{\label{tour-path}An illustration of the tour path}\label{fig:tour-path}
\end{figure}

\newpage

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/toy-search-1.pdf}
\caption{\label{toy-search}A comparison of plotting the same search
points with different plot designs. In the upper plot, points from the
last three iterations span the vast majority of the plotting space
leaving the first few iterations being squeezed together whilst the
lower plot spaces each iteration evenly and presents summary information
easy to read.}
\end{figure}

\newpage

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/toy-interp-1.pdf}
\caption{\label{toy-interp} This is a toy example of plotting
interpolated points}
\end{figure}

\newpage

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/toy-pca-1.pdf}
\caption{\label{toy-pca}PCA plot of search geodesic colouring by info
allows for better understanding of each stage in the geodesic search}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/toy-tour-1.pdf}
\caption{\label{toy-tour}A selected number of frames from the tour
animation for viewing the 5D space of all the projection bases. The
second frame on the top row view the space from a direction that is
close to the one in PCA plot. The tour animation allows for a more
holistic view of the full space in high dimensions from different
angles.}
\end{figure}

\newpage

@@

For a particular index function, the best algorithm to optimise is
related to the character of the index and the data. If the index
function is smooth and has single maximum, all of the three algorithms
introduced above can find maximum. When multiple optima are presented,
\texttt{search\_better} may stuck in the local maximum. When the index
function is non-smooth, \texttt{search\_geodesic} may even fail to find
the maximum. In this section, examples will be presented to outline how
the diagnostic plots can be used to compare the performance of the
algorithms in different scenarios.

@@

Random variables with different structures has been simulated and the
distribution of each is presented in Equation \ref{eq:sim-norm} to
\ref{eq:sim-x7}. Variable \texttt{x1}, \texttt{x8}, \texttt{x9} and
\texttt{x10} are normal distributed with zero mean and unit variance and
\texttt{x2} to \texttt{x7} are mixture of normal distributions with
different weights and locations. Figure \ref{origin-data} plots the
histogram of each variable except \texttt{x3}. The mixture variables are
scaled to have unit variance before running the projection pursuit.

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/origin-data-1.pdf}
\caption{\label{origin-data} The histogram of simulated selected
variables.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/interruption-1.pdf}
\caption{\label{interruption}Trace plots of the interpolated basis with
and without interruption. The interruption stops the interpolation when
the index value starts to decrease at id = 60. The implementation of the
interuption finds an ending basis with higher index value using fewer
steps.}
\end{figure}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{123456}\NormalTok{)}
\NormalTok{holes_2d_geo <-}\StringTok{ }\KeywordTok{animate_xy}\NormalTok{(data_mult[,}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{, }\DecValTok{7}\OperatorTok{:}\DecValTok{10}\NormalTok{)],}\DataTypeTok{tour_path =} 
                             \KeywordTok{guided_tour}\NormalTok{(}\KeywordTok{holes}\NormalTok{(), }\DataTypeTok{d =} \DecValTok{2}\NormalTok{, }
                                         \DataTypeTok{search_f =}\NormalTok{ tourr}\OperatorTok{:::}\NormalTok{search_geodesic),}
                           \DataTypeTok{rescale =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{TRUE}\NormalTok{)}

\NormalTok{last_basis <-}\StringTok{ }\NormalTok{holes_2d_geo }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(info }\OperatorTok{==}\StringTok{ "interpolation"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{tail}\NormalTok{(}\DecValTok{1}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{pull}\NormalTok{(basis) }\OperatorTok{%>%}\StringTok{ }\NormalTok{.[[}\DecValTok{1}\NormalTok{]]}

\KeywordTok{set.seed}\NormalTok{(}\DecValTok{123456}\NormalTok{)}
\NormalTok{holes_2d_geo_polish <-}\StringTok{ }\KeywordTok{animate_xy}\NormalTok{(data_mult[,}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{, }\DecValTok{7}\OperatorTok{:}\DecValTok{10}\NormalTok{)], }\DataTypeTok{tour_path =} 
                                    \KeywordTok{guided_tour}\NormalTok{(}\KeywordTok{holes}\NormalTok{(), }\DataTypeTok{d =} \DecValTok{2}\NormalTok{, }
                                                \DataTypeTok{search_f =}\NormalTok{ tourr}\OperatorTok{:::}\NormalTok{search_polish),}
                                  \DataTypeTok{rescale =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{TRUE}\NormalTok{, }
                                  \DataTypeTok{start =}\NormalTok{ last_basis)}
\end{Highlighting}
\end{Shaded}

A slight variation of the plot definition due to the addition of
polishing points is as follows:

  \begin{itemize}
  \tightlist
  \item
    data: filter the observations with \(S = \text{interpolation}\);
    bind the data object from optimisation and interpolation and form
    polishing; mutate \(t\) to be the row number of the binded tibble.
  \item
    x: \(t\) is mapped to the x-axis
  \item
    y: \(I\) is mapped to the y-axis
  \item
    colour: \(V\) is mapped to the colour aesthetic
  \end{itemize}
\item
  Layer 2: line geom
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

Again using the same data, we are interested to compare the effect of
different \texttt{max.tries} in the 2D projection setting.
\texttt{max.tries} is a hyperparameter that controls the maximum number
of try before the search ends. The default value of 25 is suitable for
1D projection while we suspect it may not be a good option for the 2D
case and hence want to compare it with an alternative, 500. As shown in
Figure \ref{trace-compare}, both trials attain the same index value
after polishing while the small \texttt{max.tries} of 25 is not
sufficient for \texttt{search\_better} to find its global maximum and we
will need to adjust the \texttt{max.tries} argument for the search to
sufficiently explore the parameter space.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/polish-1.pdf}
\caption{\label{trace-compare}Breakdown of index value when using
different max.tries in search better in conjunction with search polish.
Both attain the same final index value after the polishing while using a
\texttt{max.tries} of 25 is not sufficient to find the true maximum.}
\end{figure}

@@

{[}add example from geozoo{]}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/kol-cdf-1.pdf}
\caption{\label{kol-cdf}Comparison of two different searching methods:
search\_geodesic and search\_better on 1D projection problem for a
noisier index: kol\_cdf. The geodesic search rely heavily on the
polishing step to find the final index value while search better works
well.}
\end{figure}

@@

Using different seeds in \texttt{search\_better} allows us to find both
local maxima d as in Figure \ref{1d-2var-different-seeds}. Comparing the
maximum of both, we can see that the global maximum happens when
\texttt{x2} is found. It is natural to ask then if there is an algorithm
that can find the global maximum without trying on different seeds?
\texttt{search\_better\_random} manages to do it via a
Metropolis-hasting random search as shown in Figure
\ref{1d-2var-better-random}, although at a higher cost of number of
points to evaluate.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/1d-2var-different-seeds-1.pdf}
\caption{\label{1d-2var-different-seeds}The trace plot search better in
a 1D projection problem with two informative variables using different
seeds (without polishing). Since there are two informative variables,
setting different value for seed will lead search better to find either
of the local maximum.}
\end{figure}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/1d-2var-better-random-1.pdf}
\caption{\label{1d-2var-better-random}Using search better random for the
problem above will result in finding the global maximum but much larger
number of iteration is needed.}
\end{figure}

We can also plot the searching points of all three algorithms in the
searching space and explore their relative position against each other
using principal components. As shown in Figure
\ref{1d-2var-explore-proj-pca}, the bases from better1 and better2 only
search a proportion of the searching space while better\_random produces
a more exhaustive search. The large overlapping of better1 and
better\_random is explained by the fact that both algorithms find x2 in
the end.

>>>>

{\centering \includegraphics[width=0.6\linewidth,height=0.4\textheight]{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figures/tour_path_keynote/tour_path_keynote.001}

\caption{\label{tour-path}An illustration of the tour path}\label{fig:tour-path}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/toy-search-1.pdf}
\caption{\label{fig:toy-search}\label{toy-search}A comparison of plotting the same search points with different plot designs. In the upper plot, points from the last three iterations span the vast majority of the plotting space leaving the first few iterations being squeezed together whilst the lower plot spaces each iteration evenly and presents summary information easy to read.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/toy-interp-1.pdf}
\caption{\label{fig:toy-interp}\label{toy-interp} This is a toy example of plotting interpolated points}
\end{figure}

~

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/toy-pca-1.pdf}
\caption{\label{fig:toy-pca}\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/toy-tour-1.pdf}
\caption{\label{fig:toy-tour}\label{toy-tour}A selected number of frames from the tour animation for viewing the 5D space of all the projection bases. The second frame on the top row view the space from a direction that is close to the one in PCA plot. The tour animation allows for a more holistic view of the full space in high dimensions from different angles.}
\end{figure}

@@

For a particular index function, the best algorithm to optimise relates to the character of the index and the data. If the index function is smooth and has single maximum, all of the three algorithms introduced above can find the maximum. When multiple optima are presented, \texttt{search\_better} may stuck in the local maximum and in the case where the index function is non-smooth, \texttt{search\_geodesic} may even fail to find the maximum. In this section, examples will be presented to outline how the diagnostic plots can be used to compare the performance of the algorithms in different scenarios.

@@

Random variables with different structures has been simulated and the distribution of each is presented in Equation \ref{eq:sim-norm} to \ref{eq:sim-x7}. Variable \texttt{x1}, \texttt{x8}, \texttt{x9} and \texttt{x10} are normal distributed with zero mean and unit variance and \texttt{x2} to \texttt{x7} are mixtures of normal distributions with varied weights and locations. The mixture variables have been scaled to have an overall unit variance before running the projection pursuit.

@@

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/interruption-1.pdf}
\caption{\label{fig:interruption}\label{interruption}Trace plots of the interpolated basis with and without interruption. The interruption stops the interpolation when the index value starts to decrease at id = 60. The implementation of the interuption finds an ending basis with higher index value using fewer steps.}
\end{figure}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(tourr)}
\KeywordTok{library}\NormalTok{(ferrn)}

\KeywordTok{set.seed}\NormalTok{(}\DecValTok{123456}\NormalTok{)}
\NormalTok{holes_2d_geo <-}\StringTok{ }
\StringTok{  }\KeywordTok{animate_xy}\NormalTok{(boa6,}
             \DataTypeTok{tour_path =} \KeywordTok{guided_tour}\NormalTok{(}\KeywordTok{holes}\NormalTok{(), }\DataTypeTok{d =} \DecValTok{2}\NormalTok{, }
                                     \DataTypeTok{search_f =}\NormalTok{ search_geodesic),}
             \DataTypeTok{rescale =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{TRUE}\NormalTok{)}

\NormalTok{last_basis <-}\StringTok{ }\KeywordTok{get_best}\NormalTok{(holes_2d_geo) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{pull}\NormalTok{(basis) }\OperatorTok{%>%}\StringTok{ }\NormalTok{.[[}\DecValTok{1}\NormalTok{]]}

\KeywordTok{set.seed}\NormalTok{(}\DecValTok{123456}\NormalTok{)}
\NormalTok{holes_2d_geo_polish <-}\StringTok{ }
\StringTok{  }\KeywordTok{animate_xy}\NormalTok{(boa6, }
             \DataTypeTok{tour_path =} \KeywordTok{guided_tour}\NormalTok{(}\KeywordTok{holes}\NormalTok{(), }\DataTypeTok{d =} \DecValTok{2}\NormalTok{, }
                                     \DataTypeTok{search_f =}\NormalTok{ search_polish),}
             \DataTypeTok{rescale =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{start =}\NormalTok{ last_basis)}
\end{Highlighting}
\end{Shaded}

Checking the final projection before and after applying the polish as in Figure \ref{geo-polish}, we can see that \ldots{}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/geo-polish-1.pdf}
\caption{\label{fig:geo-polish}\label{geo-polish}On the left is the final projection from search geodesic while on the right is the polished result. Polishing helps to sharpen the edges of the clusters}
\end{figure}

@@

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/kol-cdf-1.pdf}
\caption{\label{fig:kol-cdf}\label{kol-cdf}Comparison of two different searching methods: search\_geodesic and search\_better on 1D projection problem for a noisier index: kol\_cdf. The geodesic search rely heavily on the polishing step to find the final index value while search better works well.}
\end{figure}

@@

Using different seeds in \texttt{search\_better} allows us to find both local maxima d as in Figure \ref{1d-2var-different-seeds}. Comparing the maximum of both, we can see that the global maximum happens when \texttt{x2} is found. It is natural to ask then if there is an algorithm that can find the global maximum without trying on different seeds? \texttt{search\_better\_random} manages to do it via a Metropolis-hasting random search as shown in Figure \ref{1d-2var-better-random}, although at a higher cost of number of points to evaluate. In Figure \ref{kol-pca},
sing search better random for the problem above will result in finding the global maximum but much larger number of iteration is needed. We can also plot the searching points of all three algorithms in the searching space and explore their relative position against each other using principal components. As shown in Figure \ref{1d-2var-explore-proj-pca}, the bases from better1 and better2 only search a proportion of the searching space while better\_random produces a more exhaustive search. The large overlapping of better1 and better\_random is explained by the fact that both algorithms find x2 in the end.
The trace plot search better in a 1D projection problem with two informative variables using different seeds (without polishing). Since there are two informative variables, setting different value for seed will lead search better to find either of the local maximum.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/kol-pca-1.pdf}
\caption{\label{fig:kol-pca}\label{kol-pca}THis is the caption}
\end{figure}
